---
title: "Check depth in OBIS data"
author: "Yi-Ming Gan"
date: "`r Sys.Date()`"
output: github_document
---

## Check records with depth <= 0 meters


```{r check depth record, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(httr)
library(robis)

url <- "https://api.obis.org/"
r = GET(url, path = "v3/occurrence", query = list(enddepth = 0))
content(r)$total
```

There are many records with negative depth. How many are they?

```{r depth occ count, warning=FALSE, message=FALSE}
# variables
end_depth = 0  # in meters
max_end_depth = -11000

# initialize tibble with end_depth at 100m interval
occ_df <- tibble(end_depth = seq(0, max_end_depth, -100))

# function to get occurrence record count per end_depth
get_occ_count_with_end_depth <- function(end_depth){
  url <- "https://api.obis.org/"
  r <- GET(url, path = "v3/occurrence", query = list(enddepth = end_depth))
  count <- content(r)$total  # get the count of records from query content
  return(count)
}

# create columns of record count for each of end_depth
occ_df <- occ_df %>%
  rowwise() %>%
  mutate(
    count = get_occ_count_with_end_depth(end_depth)
    )

occ_df

# plot them into bar chart
ggplot(occ_df, aes(x = end_depth, y = count)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(trans="log10")
```

## What are taxa of these records where depth < 0?

Since the web service takes integer for `enddepth` and 0 is a valid depth value. Attempt query for facets with taxon `class` with enddepth = -1.

```{r taxon class negative depth, warning=FALSE, message=FALSE}
# GET request with parameters
url <- "https://api.obis.org/"
r <- GET(url, path = "v3/facet", query = list(facets = "class", enddepth = -1))
results <- content(r)$results$class

# create a tibble of each taxonomic class with the count
class_negative_depth <- tibble(
  class = map_chr(results, "key"),
  count = map_int(results, "records")
) %>% arrange(desc(count))

class_negative_depth

ggplot(class_negative_depth, aes(x = count, y = class)) + 
  geom_col(stat = "identity")

```

Aside from birds (Aves) where the records should not use `minimumDepthInMeters` and `maximumDepthInMeters` to document the distance above surface, the rest of the records with negative depth seems to be misunderstanding of the terms definition.


